### 2022-09-28 code/0_preprocessing/0_preprocessingVN.Rmd

---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("tidyjson")
library("gtools")
library("dplyr")
library(lme4)
library(digest)
library(lmerTest)
```

### Getting VN data

```{r import data, warning=FALSE, eval=F}
# read in raw data files
#folder_path = "~/Triad/data_expt/data_V/"
all_files <- list.files(path=folder_path, pattern="")

split_responses <- function(responses) {
  responses <- as.character(responses)
  response_list <- unlist(strsplit(responses, ',|:'))
  response_list <- response_list[seq(2, length(response_list), 2)]
  response_list <- gsub("}", "", response_list)
  response_list <- gsub("\"", "", response_list)
  return(response_list)
}

colnames <- c("subject", "trial_index", "rt", "trial_type", "stim_type", "cue", "top_opt", "bottom_opt", "responses", "variable_type")

df <- data.frame()
#reading each file within the range and append them to create one file
for (f in all_files){
  file_path = paste(folder_path, f, sep="")
  result <- read_json(path = file_path, format = "infer")     # read the file
  temp <- data.frame()
  for (page in result[[2]][[1]]) { #iterating through each trial
    #if (page["subject"] %in% exclude_subjects) { next }
    if (page["trial_type"] == "survey-multi-choice") { #only triads

       trials <- matrix(NA, nrow = length(page["options"][[1]]), ncol = length(colnames), 
                        dimnames = list(NULL, colnames))
       trials <- as.data.frame(trials)

       trials$top_opt <- lapply(page["options"][[1]], `[[`, 1)
       trials$bottom_opt <- lapply(page["options"][[1]], `[[`, 2)
       trials$stim_type <- page["stimulus_type"][[1]]
       trials$cue <- page["prompt"][[1]]

       trials$responses <- split_responses(page["responses"])

       trials$subject <- page["subject"]
       trials$trial_index <- page["trial_index"]
       trials$rt <- page["rt"]
       trials$trial_type <- page["trial_type"]

    } else {
      trials <- as.data.frame(page)
    }
    temp <- smartbind(temp, select(trials, one_of(colnames)))
  }
  df <- smartbind(df, select(temp, one_of(colnames)))
}

anonymise <- function(data, cols_to_anon, algo = "sha256")
{
  if(!require(digest)) stop("digest package is required")
  to_anon <- subset(data, select = cols_to_anon)
  unname(apply(to_anon, 1, digest, algo = algo))
}

cols_to_anon <- c("subject")
df$subject <- anonymise(df, cols_to_anon)
```

```{r recoded a cue due to typo}
df_raw <- df

df <- df_raw %>%
  mutate(responses = ifelse(responses == "búa ", "búa", responses)) %>%
  mutate(top_opt = ifelse(top_opt == "búa ", "búa", top_opt)) %>%
  mutate(bottom_opt = ifelse(bottom_opt == "búa ", "búa", bottom_opt))

write.csv(df, "data_V_0530.csv")
```