### 2022-09-28 code/1_exclusions/1_exclusionsUS.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r loading data E, warning=FALSE, include=F}
df.resp_E <- read.csv("../../data/responding/data_US.csv") %>%
  select(-X)

# check for participants who completed the entire experiment. Because the experiment is
# quite long, some participants decide to come back afterwards, so there are a few files
# that are recorded as incomplete. 
df.ppts_finished_E <- df.resp_E %>% 
  group_by(subject) %>%
  summarize(finished = ifelse("demo-final" %in% unique(variable_type), "yes", "no")) %>%
  filter(finished == "yes")

df.resp_E <- df.resp_E %>%
  filter(subject %in% df.ppts_finished_E$subject)

df.ppts_att_E <- df.resp_E %>%
  filter(stim_type == "attention check") %>%
  mutate(correct_response = strsplit(cue, "Choose ")) %>%
  mutate(correct_response = sapply(correct_response, "[[", 2)) %>%
  mutate(pass_att_check = ifelse(responses == correct_response, 1, 0)) %>%
  group_by(subject) %>%
  summarize(pass_att_check_percent = mean(pass_att_check)) %>%
  filter(pass_att_check_percent == 1.0)

df.resp_E <- df.resp_E %>%
  filter(subject %in% df.ppts_att_E$subject) 
```

```{r exclusion E, include = F, warning = F}
#non-native speaker
split_responses_demog <- function(responses) {
  responses <- as.character(responses)
  response_list <- unlist(strsplit(responses, ',|:'))
  response_list <- response_list[seq(1, length(response_list))]
  response_list <- gsub("}", "", response_list)
  response_list <- gsub("\"", "", response_list)
  return(response_list)
}

NATIVE_LANG = 2
get_native_lang <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  if(responses[NATIVE_LANG-1] != "{firstlanguage") 
    {warnings("cannot find native language answer")}
  else
    {return (responses[NATIVE_LANG])}
})

df.ppts_non_native_lang_E <- df.resp_E %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_language") %>%
  mutate(native_speaker = get_native_lang(responses)) %>%
  filter(native_speaker == "No")

ppts_non_native_lang_E = unique(df.ppts_non_native_lang_E$subject)#18
percent_dropped = length(ppts_non_native_lang_E) / length(unique(df.resp_E$subject))
percent_dropped #0.1034482775

#fluent in Vietnamese & Mandarin
SPEAKER = 2
SPEAKING = 23
UNDERSTANDING = 46

get_speaker <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[SPEAKER])
})

get_fluency_speaking <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[SPEAKING])
})

get_fluency_understanding <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[UNDERSTANDING])
})

df.lang_vi_E <- df.resp_E %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_language_vi") %>%
  mutate(vi_speaker = get_speaker(responses))

df.lang_vi_fluency_E <- df.resp_E %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_conditional_language_vn_fluency") %>%
  mutate(vi_fluency_sp = as.numeric(get_fluency_speaking(responses))) %>%
  mutate(vi_fluency_ud = as.numeric(get_fluency_understanding(responses)))

df.lang_vi_E <- full_join(df.lang_vi_E, 
                          df.lang_vi_fluency_E, 
                          by = "subject") %>%
  select(subject, vi_speaker, vi_fluency_sp, vi_fluency_ud) %>%
  filter(vi_speaker == "Yes",
         vi_fluency_sp > 4, 
         vi_fluency_ud > 4)

ppts_vi_speaker_E = unique(df.lang_vi_E$subject) #4
percent_dropped = length(ppts_vi_speaker_E) / length(unique(df.resp_E$subject))
percent_dropped #0.02298851

df.lang_zh_E <- df.resp_E %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_language_cn") %>%
  mutate(zh_speaker = get_speaker(responses))

df.lang_zh_fluency_E <- df.resp_E %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_conditional_language_cn_fluency") %>%
  mutate(zh_fluency_sp = as.numeric(get_fluency_speaking(responses))) %>%
  mutate(zh_fluency_ud = as.numeric(get_fluency_understanding(responses)))

df.lang_zh_E <- full_join(df.lang_zh_E, 
                          df.lang_zh_fluency_E, 
                          by = "subject") %>%
  select(subject, zh_speaker, zh_fluency_sp, zh_fluency_ud) %>%
  filter(zh_speaker == "Yes",
         zh_fluency_sp > 4, 
         zh_fluency_ud > 4)

ppts_zh_speaker_E = unique(df.lang_zh_E$subject) #30
percent_dropped = length(ppts_zh_speaker_E) / length(unique(df.resp_E$subject))
percent_dropped #0.1742424

#have lived outside the sample country for more than 2 years
#have travelled extensively (more than 6 international experiences of 2 days or longer) 
YEARSABROAD = 2
OVERSEAEXP = 4

get_lived_abroad <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[YEARSABROAD])
})

get_oversea_exp <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[OVERSEAEXP])
})

df.abroad_E <- df.resp_E %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_oversea_experience") %>%
  mutate(lived_abroad = get_lived_abroad(responses)) %>%
  mutate(oversea_exp = get_oversea_exp(responses))

df.lived_abroad_E <- df.abroad_E %>%
  filter(lived_abroad == "Yes")

ppts_lived_abroad_E = unique(df.lived_abroad_E$subject) #29
percent_dropped = length(ppts_lived_abroad_E) / length(unique(df.resp_E$subject))
percent_dropped #0.1666667

df.oversea_exp_E <- df.abroad_E %>%
  filter(oversea_exp == "Six or more experiences")

ppts_oversea_exp_E = unique(df.oversea_exp_E$subject) #77
percent_dropped = length(ppts_oversea_exp_E) / length(unique(df.resp_E$subject))
percent_dropped #0.44 # will not use this criterion

df.resp_E <- df.resp_E %>%
  filter(!(subject %in% ppts_non_native_lang_E),
         !(subject %in% ppts_vi_speaker_E), 
         !(subject %in% ppts_zh_speaker_E),
         !(subject %in% ppts_lived_abroad_E))

length(unique(df.resp_E$subject)) #132
```

```{r}
write.csv(df.resp_E, "../../data/dataUS_excl.csv")
```