### 2022-09-28 code/1_exclusions/1_exclusionsVN.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r loading data V, include=F}
df.resp_V <- read.csv("../../data/responding/data_VN.csv") %>%
  select(-X)

# check for participants who completed the entire experiment. Because the experiment is
# quite long, some participants decide to come back afterwards, so there are a few files
# that are recorded as incomplete. 
df.ppts_finished_V <- df.resp_V %>% 
  group_by(subject) %>%
  summarize(finished = ifelse("demo-final" %in% unique(variable_type), "yes", "no")) %>%
  filter(finished == "yes")

df.resp_V <- df.resp_V %>%
  filter(subject %in% df.ppts_finished_V$subject)

df.ppts_att_V <- df.resp_V %>%
  filter(stim_type == "attention check") %>%
  mutate(correct_response = strsplit(cue, "Chọn ")) %>%
  mutate(correct_response = sapply(correct_response, "[[", 2)) %>%
  mutate(pass_att_check = ifelse(responses == correct_response, 1, 0)) %>%
  group_by(subject) %>%
  summarize(pass_att_check_percent = mean(pass_att_check)) %>%
  filter(pass_att_check_percent == 1.0)

df.resp_V <- df.resp_V %>%
  filter(subject %in% df.ppts_att_V$subject) #73
```

```{r exclusion V, include = F}
#non-native speaker
split_responses_demog <- function(responses) {
  responses <- as.character(responses)
  response_list <- unlist(strsplit(responses, ',|:'))
  response_list <- response_list[seq(1, length(response_list))]
  response_list <- gsub("}", "", response_list)
  response_list <- gsub("\"", "", response_list)
  return(response_list)
}

NATIVE_LANG = 2
get_native_lang <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  if(responses[NATIVE_LANG-1] != "{firstlanguage") 
    {warnings("cannot find native language answer")}
  else
    {return (responses[NATIVE_LANG])}
})

df.ppts_non_native_lang_V <- df.resp_V %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_language") %>%
  mutate(native_speaker = get_native_lang(responses)) %>%
  filter(native_speaker == "Không")

ppts_non_native_lang_V = unique(df.ppts_non_native_lang_V$subject)
percent_dropped = length(ppts_non_native_lang_V) / length(unique(df.resp_V$subject))
percent_dropped #0

#fluent in English
SPEAKER = 2
SPEAKING = 23
UNDERSTANDING = 46

get_speaker <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[SPEAKER])
})

get_fluency_speaking <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[SPEAKING])
})

get_fluency_understanding <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[UNDERSTANDING])
})

df.lang_en_V <- df.resp_V %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_language_en") %>%
  mutate(en_speaker = get_speaker(responses))

df.lang_en_fluency_V <- df.resp_V %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_conditional_language_en_fluency") %>%
  mutate(en_fluency_sp = as.numeric(get_fluency_speaking(responses))) %>%
  mutate(en_fluency_ud = as.numeric(get_fluency_understanding(responses)))

df.lang_en_V <- full_join(df.lang_en_V, 
                          df.lang_en_fluency_V, 
                          by = "subject") %>%
  select(subject, en_speaker, en_fluency_sp, en_fluency_ud) %>%
  filter(en_speaker == "Có",
         en_fluency_sp > 4, 
         en_fluency_ud > 4)

ppts_en_speaker_V = unique(df.lang_en_V$subject) #4
percent_dropped = length(ppts_en_speaker_V) / length(unique(df.resp_V$subject))
percent_dropped #0.6986301 # do not use this criterion

df.lang_zh_V <- df.resp_V %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_language_cn") %>%
  mutate(zh_speaker = get_speaker(responses))

df.lang_zh_fluency_V <- df.resp_V %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_conditional_language_cn_fluency") %>%
  mutate(zh_fluency_sp = as.numeric(get_fluency_speaking(responses))) %>%
  mutate(zh_fluency_ud = as.numeric(get_fluency_understanding(responses)))

df.lang_zh_V <- full_join(df.lang_zh_V, 
                          df.lang_zh_fluency_V, 
                          by = "subject") %>%
  select(subject, zh_speaker, zh_fluency_sp, zh_fluency_ud) %>%
  filter(zh_speaker == "Có",
         zh_fluency_sp > 4, 
         zh_fluency_ud > 4)

ppts_zh_speaker_V = unique(df.lang_zh_V$subject) #4
percent_dropped = length(ppts_zh_speaker_V) / length(unique(df.resp_V$subject))
percent_dropped #0.06557377

#have lived outside the sample country for more than 2 years
#have travelled extensively (more than 6 international experiences of 2 days or longer) 
YEARSABROAD = 2
OVERSEAEXP = 4

get_lived_abroad <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[YEARSABROAD])
})

get_oversea_exp <- Vectorize(function(responses) {
  responses <- split_responses_demog(responses)
  return (responses[OVERSEAEXP])
})

df.abroad_V <- df.resp_V %>%
  select(subject, responses, variable_type) %>%
  filter(variable_type == "demog_oversea_experience") %>%
  mutate(lived_abroad = get_lived_abroad(responses)) %>%
  mutate(oversea_exp = get_oversea_exp(responses))

df.lived_abroad_V <- df.abroad_V %>%
  filter(lived_abroad == "Có")

ppts_lived_abroad_V = unique(df.lived_abroad_V$subject)
percent_dropped = length(ppts_lived_abroad_V) / length(unique(df.resp_V$subject))
percent_dropped #0.1369863

df.oversea_exp_V <- df.abroad_V %>%
  filter(oversea_exp == "6 lần trở lên")

ppts_oversea_exp_V = unique(df.oversea_exp_V$subject) #77
percent_dropped = length(ppts_oversea_exp_V) / length(unique(df.resp_V$subject))
percent_dropped #0.1506849

df.resp_V <- df.resp_V %>%
  filter(!(subject %in% ppts_non_native_lang_V),
         !(subject %in% ppts_oversea_exp_V), 
         !(subject %in% ppts_zh_speaker_V),
         !(subject %in% ppts_lived_abroad_V))

length(unique(df.resp_V$subject)) 
```

```{r}
write.csv(df.resp_V, "../../data/dataVN_excl.csv")
```