### 2022-11-30 code/0_demographic/0_demographic.Rmd
added a lot of analysis. prepared to move to coglink_cogsci

---
title: "Demographics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("tidyjson")
library("gtools")
library("dplyr")
library(lme4)
library(digest)
library(lmerTest)
```

```{r}
source("extract_demog.r")
```
### Getting raw data

```{r}
df.US <- read.csv("../../data/responding/data_US.csv") %>%
  select(-X)

df.CN <- read.csv("../../data/responding/data_CN.csv") %>%
  select(-X)

df.VN <- read.csv("../../data/responding/data_VN.csv") %>%
  select(-X)
```

```{r warning=F}
df.US_demog <- extract_demog(df.US, "US")
df.VN_demog <- extract_demog(df.VN, "VN")
df.CN_demog <- extract_demog(df.CN, "CN")
df.demog <- full_join(df.US_demog, df.VN_demog) %>%
  full_join(df.CN_demog)
write.csv(df.demog, "../../data/responding/data_demog.csv")
```