### 2022-11-30 code/2_combine/2_combine_assoc.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")
```

```{r get association strengths and triad information}
df.assoc = read_delim("../../data/assoc_strengths.csv",delim='\t',quote = '',escape_backslash = F,escape_double = F) %>%
  select(cue, response, R123.Strength) %>%
  rename(assoc_strength = R123.Strength)
df.triad_E = read.csv("../../data/corpus/corpusE.csv") %>% 
  select(Cue_renamed, Cue, Tax.match, Theme.match) %>%
  rename(triad = Cue_renamed, 
         cue = Cue, 
         tax_match_E = Tax.match, 
         theme_match_E = Theme.match)

df.filler_E = read.csv("../../data/corpus/corpusEMV_filler.csv") %>% 
  select(Cue_renamed, Cue, Tax.match, Theme.match) %>%
  rename(triad = Cue_renamed, 
         cue = Cue, 
         tax_match_E = Tax.match, 
         theme_match_E = Theme.match)
```

```{r combine association strength and triad information}
df.triad_assoc <- left_join(df.triad_E, df.assoc,
                            by = c('cue'='cue', 'tax_match_E'='response')) %>%
  rename(tax_match_assoc_strength_E = assoc_strength)

df.triad_assoc <- left_join(df.triad_assoc, df.assoc,
                            by = c('cue'='cue', 'theme_match_E'='response')) %>%
  rename(theme_match_assoc_strength_E = assoc_strength)

#write.csv(df.triad_assoc, "../../data/corpus/corpusE_assoc_strength.csv", row.names = FALSE)
```

```{r}
triad_omit <- df.triad_assoc %>%
  filter(is.na(tax_match_assoc_strength_E) & is.na(theme_match_assoc_strength_E))
#write.csv(triad_omit$triad, "../../data/triads_omit_assoc.csv", row.names = FALSE)
```

```{r combine association strength data with responding data}
df = read.csv("../../data/data_USCNVN_ENZHVI.csv") %>%
  select(-X)

#df.triad_assoc <- read.csv("../../data/corpus/corpusE_assoc_strength.csv") %>%
#  select(-X)

df <- left_join(df, df.triad_assoc, 
                by = c("cue", "triad", "tax_match_E", "theme_match_E"))

write.csv(df, "../../data/data_USCNVN_ENZHVI_assoc_strength.csv", row.names = F)
```