### 2022-11-30 code/3_analysis/3_analysis_compare_freq_cosine.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")
library("Hmisc")
library("dplyr")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())

r_refs("r-references.bib")
r_refs("references/packages.bib")

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r read in data}
df <- read.csv("../../data/data_USCNVN_ENZHVI.csv")
triads_omit_freq <- read.csv("../../data/triads_omit.csv") %>%
  pull(x)
```

# Comparing between 2 models for US repsonding
```{r US responding predicted by each corpus model raw co-occurrences, warning=FALSE}
# English 
fit.EN_US = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", 
                               !(near(theme_frequency_E, .Machine$double.eps) & 
                                  near(tax_frequency_E, .Machine$double.eps))), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)

fit.EN_US.anova

fit.EN_US_cos = glmer(responses_theme ~ theme_cosine_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", 
                               !(near(theme_match_frequency_E, .Machine$double.eps) & 
                                  near(tax_match_frequency_E, .Machine$double.eps))),

          family="binomial")
summary(fit.EN_US_cos)
fit.EN_US_cos.anova = Anova(fit.EN_US_cos, type = 3)
fit.EN_US_cos.anova

fit.EN_US_compare_freq_cos = anova(fit.EN_US, fit.EN_US_cos, type = 3)
fit.EN_US_compare_freq_cos

```

# Comparing between 2 models for VN responding
```{r US responding predicted by each corpus model raw co-occurrences, warning=FALSE}
# Vietnamese 
fit.VI_VN = glmer(responses_theme ~ theme_freq_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", 
                               !(near(theme_frequency_V, .Machine$double.eps) & 
                                  near(tax_frequency_V, .Machine$double.eps))), 
          family="binomial")
summary(fit.VI_VN)
fit.VI_VN.anova = Anova(fit.EN_US, type = 3)

fit.VI_VN.anova

fit.VI_VN_cos = glmer(responses_theme ~ theme_cosine_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", 
                               !(near(theme_frequency_V, .Machine$double.eps) & 
                                  near(tax_frequency_V, .Machine$double.eps))),

          family="binomial")
summary(fit.VI_VN_cos)
fit.VI_VN_cos.anova = Anova(fit.VI_VN_cos, type = 3)
fit.VI_VN_cos.anova

fit.VI_VN_compare_freq_cos = anova(fit.VI_VN, fit.VI_VN_cos, type = 3)
fit.VI_VN_compare_freq_cos

```