### 2022-11-30 code/3_analysis/3_analysis_filler.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")
library("dplyr")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())

r_refs("r-references.bib")
r_refs("references/packages.bib")

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r incorporate base freq}
df <- read.csv("../../data/data_filler_USCNVN_ENZHVI.csv")
```

```{r correlation across China and US}
df.resp_by_country <- df %>%
  group_by(cue_id, country) %>%
  summarize(emp_theme_prop = mean(responses_word1))

df.resp_by_country <- pivot_wider(data = df.resp_by_country, 
                                  names_from = country, 
                                  values_from = emp_theme_prop)
plotUSCN = ggplot(data = df.resp_by_country,
                  mapping = aes(x = US, 
                                y = China, 
                                label = cue_id)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Word1 Chosen (US)", 
        y = "Proportion Word1 Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1) + 
  theme(axis.title = element_text(size=rel(0.75)))

plotUSVN = ggplot(data = df.resp_by_country,
                  mapping = aes(x = US, 
                                y = Vietnam, 
                                label = cue_id)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Word1 Chosen (US)", 
        y = "Proportion Word1 Chosen (Vietnam)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

plotCNVN = ggplot(data = df.resp_by_country,
                  mapping = aes(x = Vietnam,
                                y = China, 
                                label = cue_id)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Word1 Chosen (Vietnam)", 
        y = "Proportion Word1 Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

corr.USCN = apa_print(cor.test(df.resp_by_country$US, 
                  df.resp_by_country$China, na.rm=T))

corr.USVN = apa_print(cor.test(df.resp_by_country$US, 
                  df.resp_by_country$Vietnam, na.rm=T))

corr.CNVN = apa_print(cor.test(df.resp_by_country$China, 
                  df.resp_by_country$Vietnam, na.rm=T))
```

```{r warning = F, fig.width = 8, fig.cap="Correlation between proportion thematic chosen by US-China participants, US-Vietnam participants, and China-Vietnam participants. All three comparisons show strong correlation, suggesting a strong signal of universality across cultural contexts regarding similarity judgments"}
plotUSCN + plotUSVN + plotCNVN +
  plot_layout(ncol = 3)
```

```{r}
df.cue_country_freq <- df %>%
  group_by(cue_id, country, word1_match_frequency_prop_E, word1_match_frequency_prop_M, word1_match_frequency_prop_V) %>%
  summarize(emp_word1_prop = mean(responses_word1))
ggplot(data = df.cue_country_freq %>% filter(country == "US"), 
        mapping = aes(x = emp_word1_prop, 
                     y = word1_match_frequency_prop_E,
                     label = cue_id)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Word1 Chosen (US)", 
        y = "Proportion Word1 Predicted (English)") + 
  theme(axis.title = element_text(size=rel(0.75)))

ggplot(data = df.cue_country_freq %>% filter(country == "Vietnam"), 
        mapping = aes(x = emp_word1_prop, 
                     y = word1_match_frequency_prop_V,
                     label = cue_id)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Word1 Chosen (VN)", 
        y = "Proportion Word1 Predicted (Vietnamese)") + 
  theme(axis.title = element_text(size=rel(0.75)))
```
# Does lexical co-occurrences for fillers in English predict US responding? 
```{r US responding predicted by each corpus model raw co-occurrences, warning=FALSE}
# English 
fit.EN_US = glmer(responses_word1 ~ word1_match_frequency_prop_E + (1 | subject) + (1 | cue), 
          data = df %>% filter(country == "US", 
                               !(near(word1_match_frequency_E, .Machine$double.eps) & 
                                  near(word2_match_frequency_E, .Machine$double.eps))), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)

fit.EN_US.anova
```
# Does lexical co-occurrences for fillers in Vietnamese predict VN responding? 
```{r VN responding predicted by each corpus model raw co-occurrences, warning=FALSE}
# Vietnamese
fit.VI_VN = glmer(responses_word1 ~ word1_match_frequency_prop_V + (1 | subject) + (1 | cue), 
          data = df %>% filter(country == "Vietnam", 
                               !(near(word1_match_frequency_V, .Machine$double.eps) & 
                                  near(word2_match_frequency_V, .Machine$double.eps))), 
          family="binomial")
summary(fit.VI_VN)
fit.VI_VN.anova = Anova(fit.VI_VN, type = 3)

fit.VI_VN.anova
```