### 2022-11-30 code/3_analysis/3_analysis_semcoding.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")
library("emmeans")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())

r_refs("r-references.bib")
r_refs("references/packages.bib")

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r combine corpus and behavior data, include = F}
source("../2_combine/2_combine.R") # include triad_omits. TODO: Find a way to also save this.
```

```{r loading in the csv with direction so we dont need to go through the previous section, include =F}
df <- read.csv("../../data/data_USCNVN_ENZHVI_direction.csv")
```

```{r loading in the csv with semantic coding, include =F}
df.sem_code <- read.csv("../../data/corpus/corpus_sem_coding.csv")

df <- df %>%
  left_join(df.sem_code, by = "triad")
```

## Do people make more thematic matches for certain semantic groups of cue?

Plotting mean response theme by country, by type of triad
```{r plotting thematic responses by cue_genre}

ggplot(data = df) + 
   geom_bar(stat = "summary", 
            fun.data = "mean_cl_boot",
            alpha = .5, 
            mapping = aes(x = cue_genre, 
                          y = responses_theme)) +
  geom_pointrange(stat = "summary", 
            fun.data = "mean_cl_boot",
            alpha = .5, 
            mapping = aes(x = cue_genre, 
                          y = responses_theme))

```

```{r plotting thematic response by cue_genre and country}
ggplot(data = df,
       mapping = aes(x = cue_genre, 
                     y = responses_theme,
                     color = country, 
                     fill = country)) + 
   stat_summary(fun.data = "mean_cl_boot", 
               geom = "pointrange") 
#color code by cue genre, spatial sep by country, plotting all triads (mean response)

```

```{r plotting thematic response by cue_genre, country and triad}
ggplot(data = df, 
       mapping = aes(x = country, 
                     y = responses_theme, 
                     label = triad, 
                     group = triad,
                     color = cue_genre)) + 
  stat_summary(fun.data = "mean_cl_boot", 
               geom = "point", 
               alpha = .5, 
               position = position_dodge(width = .2)) + 
  stat_summary(fun.data = "mean_cl_boot", 
               geom = "text", 
               position = position_dodge(width = .2), size = 2)
```

There seems to be a clear preference for more thematic matches in the 'living things' cues in comparison to the 'moveable artifacts' cue. (need to omit cue_genre = "other")

Comparing some models:
Q: SHOULD we look at a model where responses theme is predicted by genre?

### English
Base line model:

```{r US responding predicted by corpus co-occurrences, warning=FALSE}
# English 
fit.EN_US = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)
```

Adding cue genre: significant

```{r US responding, co-occurrence and cue_genre}
fit.EN_US_cue_genre = glmer(responses_theme ~ theme_freq_prop_E + cue_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")

summary(fit.EN_US_cue_genre)
Anova(fit.EN_US_cue_genre, type = 3)
```

Anova comparison: including cue genre is a better model
```{r US anova comparison, cue_genre}
fit.EN_US_cue_genre_compare = anova(fit.EN_US, fit.EN_US_cue_genre, type = 3)
fit.EN_US_cue_genre_compare
```
### Vietnamese

```{r VN responding predicted by VN corpus model, warning=FALSE}

#Vietnamese
fit.VI_VN = glmer(responses_theme ~ theme_freq_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", 
                               !(triad %in% triads_omit), 
                               !(cue_genre == "other")), 
          family="binomial")
summary(fit.VI_VN)
fit.VI_VN.anova = Anova(fit.VI_VN, type = 3)
```

Adding cue genre: NOT significant

```{r VN responding, co-occurrence + cue_genre}
fit.VI_VN_cue_genre = glmer(responses_theme ~ theme_freq_prop_V + cue_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", 
                               !(triad %in% triads_omit), 
                               !(cue_genre == "other")), 
          family="binomial")

summary(fit.VI_VN_cue_genre)
Anova(fit.VI_VN_cue_genre, type = 3)
```

Anova comparison (also not significant)
```{r VN anova comparison, cue_genre}
fit.VI_VN_cue_genre_compare = anova(fit.VI_VN, fit.VI_VN_cue_genre, type = 3)
fit.VI_VN_cue_genre_compare
```

### Mandarin
```{r CN responding predicted by co-occurrences, warning=FALSE}
#Mandarin
fit.ZH_CN = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")
summary(fit.ZH_CN)
fit.ZH_CN.anova = Anova(fit.ZH_CN, type = 3)
```

Adding cue genre: significant

```{r CN responding, co-occurrences + cue_genre}
fit.ZH_CN_cue_genre = glmer(responses_theme ~ theme_freq_prop_M + cue_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")

summary(fit.ZH_CN_cue_genre)
Anova(fit.ZH_CN_cue_genre, type = 3)
```
Anova comparison: including cue genre is a better model
```{r CN anova comparison,cue_genre}
fit.ZH_CN_cue_genre_compare = anova(fit.ZH_CN, fit.ZH_CN_cue_genre, type = 3)
fit.ZH_CN_cue_genre_compare
```

When added to the baseline model, cue genre is a significant predictor for US and China responses. But not for Vietnam.

### More general models that predict thematic response based on cue genre
NOT significant when just predicted by cue_genre.
```{r responding predicted only on cue genre, warning=FALSE}
fit.cue_genre = glmer(responses_theme ~ cue_genre + (1 | subject) + (1 | triad),
          data = df %>% filter(!(cue_genre == "other")),  
          # no need to pull out triads_omit because we're not using the raw coocurrences
          family="binomial")
summary(fit.cue_genre)
fit.cue_genre.anova = Anova(fit.cue_genre, type = 3)
fit.cue_genre.anova
```

NOT significant when predicted by country and cue_genre (only country is significant)
```{r responding predicted on cue genre and country, warning=FALSE}
fit.cue_genre_country = glmer(responses_theme ~ country + cue_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(!(cue_genre == "other")),  
          # no need to pull out triads_omit because we're not using the raw coocurrences
          family="binomial")
summary(fit.cue_genre_country)
fit.cue_genre_country.anova = Anova(fit.cue_genre_country, type = 3)
fit.cue_genre_country.anova
```

NO interaction between country and cue genre either. 
```{r responding predicted on cue genre and country, with interaction, warning=FALSE}
fit.cue_genre_country_int = glmer(responses_theme ~ country + cue_genre + country * cue_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(!(cue_genre == "other")),  
          # no need to pull out triads_omit because we're not using the raw coocurrences
          family="binomial")
summary(fit.cue_genre_country_int)
fit.cue_genre_country_int.anova = Anova(fit.cue_genre_country_int, type = 3)
fit.cue_genre_country_int.anova
```

NOT significance even with country having random slope and intercept for cue_genre.

```{r responding predicted on cue genre and country, with random intercept and slope by country, warning=FALSE}
fit.cue_genre_country_slope = glmer(responses_theme ~ country + cue_genre + (cue_genre | country) + (1 | subject) + (1 | triad), 
          data = df %>% filter(!(cue_genre == "other")),  
          # no need to pull out triads_omit because we're not using the raw co-ocurrences
          family="binomial")
summary(fit.cue_genre_country_slope)
fit.cue_genre_country_slope.anova = Anova(fit.cue_genre_country_slope, type = 3)
fit.cue_genre_country_slope.anova
```

## Do people make more thematic matches for certain semantic groups of thematic relationships?

Plotting mean response theme by country, by type of thematic match
```{r plotting thematic responses by theme_genre}
ggplot(data = df) + 
   geom_bar(stat = "summary", 
            fun.data = "mean_cl_boot",
            alpha = .5, 
            mapping = aes(x = theme_genre, 
                          y = responses_theme)) +
  geom_pointrange(stat = "summary", 
            fun.data = "mean_cl_boot",
            alpha = .5, 
            mapping = aes(x = theme_genre, 
                          y = responses_theme)) +
  geom_hline(yintercept = 0.5)

```
```{r plotting thematic responses by theme_genre and country}
ggplot(data = df) + 
   geom_bar(stat = "summary", 
            fun.data = "mean_cl_boot",
            mapping = aes(x = theme_genre, 
                          y = responses_theme, 
                          fill = country), 
            position = position_dodge(width = .7), width = .7) +
  geom_pointrange(stat = "summary", 
            fun.data = "mean_cl_boot",
            mapping = aes(x = theme_genre, 
                          y = responses_theme,
                          group = country), 
            position = position_dodge(width = .7)) +
  geom_hline(yintercept = 0.5)
#color code by cue genre, spatial sep by country, plotting all triads (mean response)

```

```{r plotting thematic response by country, triad and theme_genre}
ggplot(data = df, 
       mapping = aes(x = country, 
                     y = responses_theme, 
                     label = triad, 
                     group = triad)) + 
  geom_point(stat = "summary", 
             fun.data = "mean_cl_boot", 
             position = position_dodge(width = .7)) +
  geom_text(stat = "summary", 
            fun.data = "mean_cl_boot", 
            color = alpha('black'),
            position = position_dodge(width = .7), size = 2) +
  geom_violin(data = df %>% 
                group_by(country, theme_genre, triad) %>%
                summarize(theme_resp_triad = mean(responses_theme, na.rm = T)), 
                mapping = aes(x = country,
                              y = theme_resp_triad, 
                              group = country,
                              fill = country,
                              color = country), 
                alpha = .5) + 
  facet_grid(~theme_genre)

# geom_point(data = df %>% filter(theme_genre == 'ecology'),
#            mapping = aes(x = country,
#                    y = responses_theme,
#                    label = triad,
#                    group = triad,
#                    color = theme_genre),
#            stat = "summary",
#            fun.data = "mean_cl_boot",
#            position = position_dodge(width =.7))

```

1. All theme match semantic categories prefer thematic, except for association.
2. China is always the most thematic in all theme match semantic categories. Only country that prefers thematic across all categories. 
3. Only US & Vietnam + association prefer taxonomic. Overall, association seems to be least thematic preferring.
4. Vietnam is not significantly different from US except for ecology. 
5. Most thematic: causal (but might be driven by strong China response) and occupation (most consistent across countries to be thematic).

### English
Base line model:

```{r US responding predicted by corpus co-occurrences re-run for theme_genre, warning=FALSE}
# English 
fit.EN_US = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)
```

Adding theme genre: NOT significant (p = 0.08)

```{r US responding, co-occurrence and theme_genre}
fit.EN_US_theme_genre = glmer(responses_theme ~ theme_freq_prop_E + theme_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")

summary(fit.EN_US_theme_genre)
Anova(fit.EN_US_theme_genre, type = 3)
```

Anova comparison: including theme genre is NOT a better model
```{r US anova comparison, theme_genre}
fit.EN_US_theme_genre_compare = anova(fit.EN_US, fit.EN_US_theme_genre, type = 3)
fit.EN_US_theme_genre_compare
```
### Vietnamese

```{r VN responding predicted by VN corpus model re-run for theme_genre, warning=FALSE}

#Vietnamese
fit.VI_VN = glmer(responses_theme ~ theme_freq_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", 
                               !(triad %in% triads_omit), 
                               !(cue_genre == "other")), 
          family="binomial")
summary(fit.VI_VN)
fit.VI_VN.anova = Anova(fit.VI_VN, type = 3)
```

Adding theme genre: significant (p < 0.05)

```{r VN responding, co-occurrence + theme_genre}
fit.VI_VN_theme_genre = glmer(responses_theme ~ theme_freq_prop_V + theme_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", 
                               !(triad %in% triads_omit), 
                               !(cue_genre == "other")), 
          family="binomial")

summary(fit.VI_VN_theme_genre)
Anova(fit.VI_VN_theme_genre, type = 3)
```

Anova comparison: significant
```{r VN anova comparison, theme_genre}
fit.VI_VN_theme_genre_compare = anova(fit.VI_VN, fit.VI_VN_theme_genre, type = 3)
fit.VI_VN_theme_genre_compare
```

### Mandarin
```{r CN responding predicted by co-occurrences re-run for theme_genre, warning=FALSE}
#Mandarin
fit.ZH_CN = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")
summary(fit.ZH_CN)
fit.ZH_CN.anova = Anova(fit.ZH_CN, type = 3)
```

Adding theme genre: NOT significant (p = 0.12)

```{r CN responding, co-occurrences + theme_genre}
fit.ZH_CN_theme_genre = glmer(responses_theme ~ theme_freq_prop_M + theme_genre + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", 
                               !(triad %in% triads_omit),
                               !(cue_genre == "other")), 
          family="binomial")

summary(fit.ZH_CN_theme_genre)
Anova(fit.ZH_CN_theme_genre, type = 3)
```
Anova comparison: including theme genre is a better model
```{r CN anova comparison, theme_genre}
fit.ZH_CN_theme_genre_compare = anova(fit.ZH_CN, fit.ZH_CN_theme_genre, type = 3)
fit.ZH_CN_theme_genre_compare
```

When added to the baseline model, theme genre is a significant predictor for Vietnam response, but not for US and China!

```{r summary of semantic coding result}
text_tbl <- data.frame(
  Country = c("US", "China", "Vietnam"),
  Add_Cue_Genre = c(
    "significant", "significant", "not significant"
  ), 
  Add_Theme_Genre = c(
    "not significant", "not significant", "significant"
  )
)
kable(text_tbl)
```