### 2022-12-14 code/2_combine/2_combine_filler_assoc.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")
```

```{r get association strengths and triad information}
df.assoc = read_delim("../../data/assoc_strengths.csv",delim='\t',quote = '',escape_backslash = F,escape_double = F) %>%
  select(cue, response, R123.Strength) %>%
  rename(assoc_strength = R123.Strength)

df.filler_E = read.csv("../../data/corpus/corpusEMV_filler.csv") %>% 
  select(cue_E, word1_E, word2_E, cue_M, word1_M, word2_M, cue_V, word1_V, word2_V)
```

```{r combine association strength and triad information}
df.filler_assoc <- left_join(df.filler_E, df.assoc,
                            by = c('cue_E'='cue', 'word1_E'='response')) %>%
  rename(word1_assoc_strength_E = assoc_strength)

df.filler_assoc <- left_join(df.filler_assoc, df.assoc,
                            by = c('cue_E'='cue', 'word2_E'='response')) %>%
  rename(word2_assoc_strength_E = assoc_strength)

#write.csv(df.filler_assoc, "../../data/corpus/fillers_assoc_strength.csv", row.names = FALSE)
```

```{r}
triad_omit <- df.filler_assoc %>%
  filter(is.na(word1_assoc_strength_E) & is.na(word2_assoc_strength_E))
#write.csv(triad_omit$triad, "../../data/fillers_omit_assoc.csv", row.names = FALSE)
```

```{r combine association strength data with responding data}
df = read.csv("../../data/data_filler_USCNVN_ENZHVI.csv") %>%
  select(-X)

#df.triad_assoc <- read.csv("../../data/corpus/corpusE_assoc_strength.csv") %>%
#  select(-X)

df <- left_join(df, df.filler_assoc, 
                by = c("cue_E", "word1_E", "word2_E",
                       "cue_M", "word1_M", "word2_M", 
                       "cue_V", "word1_V", "word2_V"))

write.csv(df, "../../data/data_filler_USCNVN_ENZHVI_assoc_strength.csv", row.names = F)
```