### 2022-12-31 code/2_combine/2_combine_filler.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r loading data E, warning=FALSE, include=F}
df.resp_E <- read.csv("../../data/responding/dataUS_excl.csv") %>%
  select(-X)
```

```{r loading data V, include=F}
df.resp_V <- read.csv("../../data/responding/dataVN_excl.csv") %>%
  select(-X)
```

```{r loading data M, warning=FALSE, include=F}
df.resp_M <- read.csv("../../data/responding/dataCN_excl.csv") %>%
  select(-X)
```

```{r corpus filler just English, include = F}
df.corpus_filler <- read.csv("../../data/corpus/corpusEMV_filler.csv") %>%
  select(ends_with("_E") | ends_with("_M") | ends_with("_V")) %>%
  select(!(contains("cue_frequency") | contains("word1_frequency") | contains("word2_frequency"))) %>% # ignore base frequencies for now
  mutate(cue_id = cue_E)
```

```{r corpus - frequency, include = F}
#find triads_omit, used whenever analysis uses raw collocates 
triads_omit <- df.corpus_filler %>% 
  filter((word1_match_frequency_E == 0 & word2_match_frequency_E == 0) |
          #(word1_match_frequency_M == 0 & word2_match_frequency_M == 0) |
           (word1_match_frequency_V == 0 & word2_match_frequency_V == 0)) %>%
  pull(cue_id)
```

```{r full corpus, include=F}
#replacing any 0 raw counts with epsilon
df.corpus_filler <- df.corpus_filler %>%
    mutate(across(contains("frequency") , 
                  ~ ifelse(. == 0, .Machine$double.eps, .)))
```

```{r getting triads from the 3 languages}
#combine response and corpus data
df.filler_E <- df.resp_E %>%
  filter(stim_type == "filler",
         responses != "NA") %>%
  select(-variable_type, -trial_index, -trial_type, -stim_type) %>%
  mutate(language = "English", 
         country = "US")

df.filler_E <- left_join(df.filler_E, df.corpus_filler, 
                         by = c("cue" = "cue_E"), 
                         keep = TRUE)

df.filler_E <- df.filler_E %>%
  mutate(responses_word1 = ifelse(responses == word1_E, 1, 0))
```

```{r}
df.filler_M <- df.resp_M %>%
  filter(stim_type == "filler",
         responses != "NA") %>%
  select(-variable_type, -trial_index, -trial_type, -stim_type) %>%
  mutate(language = "Mandarin", 
         country = "China")

df.filler_M <- inner_join(df.filler_M, df.corpus_filler, 
                          by = c("cue" = "cue_M"), keep = TRUE)

df.filler_M <- df.filler_M %>%
  mutate(responses_word1 = ifelse(responses == word1_M, 1, 0))
```

```{r}
df.filler_V <- df.resp_V %>%
  filter(stim_type == "filler",
         responses != "NA") %>%
  select(-variable_type, -trial_index, -trial_type, -stim_type) %>%
  mutate(language = "Vietnamese", 
         country = "Vietnam")

df.filler_V <- inner_join(df.filler_V, df.corpus_filler, 
                          by = c("cue" = "cue_V"), keep = TRUE)

df.filler_V <- df.filler_V %>%
  mutate(responses_word1 = ifelse(responses == word1_V, 1, 0))
```

```{r}
df <- rbind(df.filler_E, df.filler_M)
df <- rbind(df, df.filler_V)

df <- df %>%
  mutate(word1_match_frequency_prop_E = 
           word1_match_frequency_E / (word1_match_frequency_E + word2_match_frequency_E),
         word1_match_frequency_prop_M = 
           word1_match_frequency_M / (word1_match_frequency_M + word2_match_frequency_M),
         word1_match_frequency_prop_V = 
           word1_match_frequency_V / (word1_match_frequency_V + word2_match_frequency_V), 
         word1_match_cosine_prop_E = 
           word1_match_cosine_E / (word1_match_cosine_E + word2_match_cosine_E),
         word1_match_cosine_prop_M = 
           word1_match_cosine_M / (word1_match_cosine_M + word2_match_cosine_M),
         word1_match_cosine_prop_V = 
           word1_match_cosine_V / (word1_match_cosine_V + word2_match_cosine_V), 
         )

write.csv(df, "../../data/data_filler_USCNVN_ENZHVI.csv", row.names = F)
write.csv(triads_omit, "../../data/fillers_omit.csv", row.names = F)
```