### 2023-01-26 code/1_exclusions_exploratory/1_exclusionsUS.Rmd

```{r setup, echo = FALSE, include = FALSE}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("patchwork")
library("effsize")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r loading data E, warning=FALSE, include=F}
df.resp_E <- read.csv("../../data/responding/data_US.csv") %>%
  select(-X)

# check for participants who completed the entire experiment. Because the experiment is
# quite long, some participants decide to come back afterwards, so there are a few files
# that are recorded as incomplete. 
df.ppts_finished_E <- df.resp_E %>% 
  group_by(subject) %>%
  summarize(finished = ifelse("demo-final" %in% unique(variable_type), "yes", "no")) %>%
  filter(finished == "yes")

df.resp_E <- df.resp_E %>%
  filter(subject %in% df.ppts_finished_E$subject)

df.ppts_att_E <- df.resp_E %>%
  filter(stim_type == "attention check") %>%
  mutate(correct_response = strsplit(cue, "Choose ")) %>%
  mutate(correct_response = sapply(correct_response, "[[", 2)) %>%
  mutate(pass_att_check = ifelse(responses == correct_response, 1, 0)) %>%
  group_by(subject) %>%
  summarize(pass_att_check_percent = mean(pass_att_check)) %>%
  filter(pass_att_check_percent >= 0.8)

df.resp_E_aft_attn <- df.resp_E %>%
  filter(subject %in% df.ppts_att_E$subject)

ppts_resp_E_aft_attn <- unique(df.resp_E_aft_attn$subject)
```

```{r}
df.demog <- read.csv("../../data/responding/data_demog.csv") %>%
  select(-X) %>%
  filter(country == "US") %>%
  filter(subject %in% ppts_resp_E_aft_attn)
```

```{r exclusion E, include = F, warning = F}
#non-native speaker
df.ppts_non_native_lang_E <- df.demog %>%
  filter(demog_question == "firstlanguage" & demog_response == "No")

ppts_non_native_lang_E = unique(df.ppts_non_native_lang_E$subject)#20
percent_dropped = length(intersect(ppts_non_native_lang_E, ppts_resp_E_aft_attn)) / length(ppts_resp_E_aft_attn)
percent_dropped #0.1034483

#fluent in Vietnamese & Mandarin
df.lang_sp_vi_E <- df.demog %>%
  filter(demog_question == "lang_sp_vi" & as.numeric(demog_response) > 4)

df.lang_ud_vi_E <- df.demog %>%
  filter(demog_question == "lang_ud_vi" & as.numeric(demog_response) > 4)

df.lang_vi_E <- inner_join(df.lang_sp_vi_E, df.lang_ud_vi_E, by = 'subject')

ppts_vi_speaker_E = unique(df.lang_vi_E$subject) #4
percent_dropped = length(intersect(ppts_vi_speaker_E, ppts_resp_E_aft_attn)) / length(ppts_resp_E_aft_attn)
percent_dropped #0.02083333

df.lang_sp_zh_E <- df.demog %>%
  filter(demog_question == "lang_sp_zh" & as.numeric(demog_response) > 4)

df.lang_ud_zh_E <- df.demog %>%
  filter(demog_question == "lang_ud_zh" & as.numeric(demog_response) > 4)

df.lang_zh_E <- inner_join(df.lang_sp_zh_E, df.lang_ud_zh_E, by = 'subject')

ppts_zh_speaker_E = unique(df.lang_zh_E$subject) #35
percent_dropped = length(intersect(ppts_zh_speaker_E, ppts_resp_E_aft_attn)) / length(ppts_resp_E_aft_attn)
percent_dropped #0.1822917

#have lived outside the sample country for more than 2 years
df.lived_abroad_E <- df.demog %>%
  filter(demog_question == "abroadexp" & demog_response == "Yes")

ppts_lived_abroad_E = unique(df.lived_abroad_E$subject) #33
percent_dropped = length(intersect(ppts_lived_abroad_E, ppts_resp_E_aft_attn)) / length(ppts_resp_E_aft_attn)
percent_dropped #0.171875

#have travelled extensively (more than 6 international experiences of 2 days or longer)
df.oversea_exp_E <- df.demog %>%
  filter(demog_question == "overseaexpnum" & demog_response == "Six or more experiences")

ppts_oversea_exp_E = unique(df.oversea_exp_E$subject) #91
percent_dropped = length(intersect(ppts_oversea_exp_E, ppts_resp_E_aft_attn)) / length(unique(df.resp_E_aft_attn$subject))
percent_dropped #0.4739583 # will not use this criterion

df.resp_E_aft_demog <- df.resp_E_aft_attn %>%
  filter(!(subject %in% ppts_non_native_lang_E),
         !(subject %in% ppts_vi_speaker_E), 
         !(subject %in% ppts_zh_speaker_E),
         !(subject %in% ppts_lived_abroad_E))

length(unique(df.resp_E_aft_demog$subject)) #119

```

```{r}
df.demog_metadata <- data.frame(
  country = c("US"),
  ppts_finished = c(length(unique(df.ppts_finished_E$subject))),
  ppts_att_check_excl = c(length(unique(df.ppts_finished_E$subject)) - length(unique(df.ppts_att_E$subject))), 
  ppts_after_att_check_excl = c(length(unique(df.ppts_att_E$subject))),
  ppts_non_native_speaker = c(paste(length(ppts_non_native_lang_E),"*")), 
  ppts_en_speaker = c("N/A"),
  ppts_vi_speaker = c(paste(length(ppts_vi_speaker_E),"*")), 
  ppts_zh_speaker = c(paste(length(ppts_zh_speaker_E),"*")), 
  ppts_lived_abroad = c(paste(length(ppts_lived_abroad_E),"*")), 
  ppts_oversea_exp = c(length(ppts_oversea_exp_E)),
  ppts_after_demog_excl = c(length(unique(df.resp_E_aft_demog$subject))))
# colnames(df.demog_metadata) <- c("Participants Finished", 
#                                  "Participants ")
```

```{r}
write.csv(df.resp_E_aft_demog, "../../data/responding/exploratory/dataUS_excl.csv")
write.csv(df.demog_metadata, "../../data/responding/exploratory/dataUS_excl_metadata.csv")
```