### 2023-02-02 code/4_paper/manuscript/manuscript.tex

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Re-examining cross-cultural similarity judgments using language
statistics}

\usepackage[utf8]{inputenc}
\usepackage[T1, T2A, T5]{fontenc}
\usepackage{kotex}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\author{{\large \bf Khuyen N. Le (knl005@ucsd.edu)}$^{1}$, {\large \bf Alexandra Carstensen (abc@ucsd.edu)}$^{1}$, \\ {\large \bf Shan Gao (shangaocog@gmail.com)$^2$}, {\large \bf Michael C. Frank (mcfrank@stanford.edu)$^3$},  \AND $^1$Department of Psychology, University of California, San Diego, $^2$Department of Psychology, University of Chicago, \\ $^3$Department of Psychology, Stanford University }

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}

\begin{document}

\maketitle

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or to ``chicken''? Speakers
of different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgments toward taxonomic relations, like cow and chicken. We measure
similarity judgments across the US, China, and Vietnam and replicate
US-China cultural differences with Vietnam intermediate. We also show
that word embedding models (fastText models for each language) are
related to judgments within each country, suggesting a possible
alternative interpretation of cross cultural differences. Perhaps
notions of similarity are similar across contexts, but the statistics of
the linguistic environment vary.

\textbf{Keywords:}
similarity; culture; language; semantics; lexical co-occurrence;
variation; US; China; Vietnam
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Many cognitive processes rely on similarity, from inference and
generalization to analogy, mathematics, and science. There is
substantial consistency in human similarity reasoning, but also
systematic variation across cultural and linguistic contexts. In
particular, there is considerable evidence showing that similarity
reasoning in East Asian cultural contexts differs from that in Western
cultures (e.g. Nisbett \& Masada, 2003). For example, in a triad task
comparing preferences for taxonomic and thematic similarity (choose two
out of three words that are most related to one another), Ji, Zhang, \&
Nisbett (2004) found that Chinese participants preferred thematic
matching to a greater extent than European Americans. In an image
version of this task, Chinese children (9-10 years old) are also more
likely to choose thematic matches compared to their American
counterparts (Chiu, 1972). This cross-cultural difference is also
observed in novel object categorization, with Chinese participants
preferring to group by family resemblance across multiple features and
Americans preferring a single-feature rule (Norenzayan, Smith, Kim, \&
Nisbett, 2002). Across tasks, East Asian participants show a preference
for thematic similarity based on causal, spatial, and temporal
relationships while Western participants are more likely to make
taxonomic matches based on the similarity of attributes, like shared
color or shape, among objects (see Markman \& Hutchinson, 1984 for a
discussion of these types of similarity).

An influential perspective within cultural psychology links these
differences in similarity judgment to tendencies toward analytic
processing in Western cultures, and holistic processing in East Asian
cultures (Nisbett, 2003). Analytic processing emphasizes rule-like
relationships predicated on objects and their properties and,
correspondingly, taxonomic similarity, while holistic processing
emphasizes relations between objects and their context and therefor,
thematic similarity. In related work, East Asian participants show a
higher level of sensitivity to context than their Western counterparts
when reproducing drawings from memory (Ji, Peng, \& Nisbett, 2000);
visually exploring naturalistic scenes (Chua, Boland, \& Nisbett, 2005);
describing scenes (Masuda \& Nisbett, 2001); and in explaining the
causes of ambiguous behaviors (Choi, Nisbett, \& Norenzayan, 1999). The
consensus interpretation in this literature ascribes cross-cultural
differences in similarity judgment to variation in the
\emph{conceptualization} of similarity -- with people from East Asian
cultures relying on a more thematic notion of similarity than
Westerners.

Alternatively, these judgments could be shaped by cross-cultural
differences in the \emph{input} to similarity judgment, that is, the
statistics of the linguistic and/or physical environment as manifest in
everyday experiences. Perhaps when faced with the triad task,
participants from all cultures follow the same process for
conceptualizing similarity, but rely on language or culture-specific
input to this process. If we observe a difference in categorization
between East Asian and Western participants, it could be that members of
both groups use the same procedure (considering similarity that is
influenced by both taxonomic and thematic relations), but that the input
to this procedure differ between cultures, with East Asian participants
exposed to more support for thematic similarity in their experience in
comparison to their Western counterparts.

\hypertarget{estimating-variation-in-experience-via-language-statistics}{%
\subsection{Estimating variation in experience via language
statistics}\label{estimating-variation-in-experience-via-language-statistics}}

To determine whether varied input to similarity judgments can in part
explain cross-cultural differences, we need a way to operationalize
variation in exposure to thematic and taxonomic relations. While
exposure to types of similarity in everyday experience is difficult to
measure, the statistics of language can provide a rough proxy. Language
statistics are useful in that they are part of the input to everyday
experience -- and indeed, may afford many of the ``experiences'' that
people have with infrequently encountered items, like cows or
helicopters -- and they provide an accessible measure. Previous work
also suggests that language statistics, such as lexical co-occurrence or
cosine distance of word embeddings, can be good predictors of similarity
reasoning. Semantic models that are constructed using lexical
co-occurrence (in comparison to annotated relations) have been shown to
perform well on predicting human judgments about similarity between word
pairs that are thematically or taxonomically related (Rohde, Gonnerman,
\& Plaut, 2006). Relatedly, a model trained on word-document
co-occurrence can predict word association and the effects of semantic
association on a variety of linguistic tasks (Griffiths, Steyvers, \&
Tenenbaum, 2007). Word embeddings like word2vec, gloVe, and fastText
have also been shown to be good predictors for similarity judgments Asr,
Zinkov, \& Jones (2018).

Our study uses cosine distances of fastText word vectors as a measure of
lexical co-occurrence.\footnote{We also carried out our analysis using
  raw lexical co-occurrences and obtained similar results.} fastText is
a system that uses lexical co-occurrence information to generate a
vector representing each word in its lexicon (Mikolov, Grave,
Bojanowski, Puhrsch, \& Joulin, 2018). Practically speaking, fastText
vectors are available for many languages, and fastText has also been
shown to be sensitive to cultural differences in word meanings: Thompson
\& Lupyan (2020) demonstrated that the distribution of semantic meaning
clusters generated by fastText trained on language-specific corpora
correlates with the cultural, historical, and geographical similarities
of these languages.

Language statistics may incorporate both cultural-specific environmental
statistics, that is, the experiential \emph{input} to similarity
judgments, and culture-specific senses of similarity, the
\emph{conceptualization} of similarity. For example, cultural features
(like farming) can lead to differences in environmental statistics
(seeing cows and grass) and these can influence language (talking more
about cows and grass). But other cultural features (like conceiving of
similarity thematically) could also cause individuals to talk
differently about the same experiences (mentioning what cows eat rather
than what other animals cows are like). Accordingly, our approach
examines the extent to which language statistics can predict
cross-cultural differences in similarity judgments with the
understanding that language statistics are likely a proxy for both
\emph{input} to and \emph{conceptualization} of similarity.

\hypertarget{the-present-study}{%
\subsection{The present study}\label{the-present-study}}

We measured taxonomic versus thematic similarity matching in a
forced-choice word triad task in three populations. Following Ji et al.
(2004), we measured preferences in the US and China. In addition, we
collected data in a novel context: Vietnam. Vietnam is a Southeast Asian
country that borders China and has historically been greatly influenced
by Chinese culture (Hui, 2002). Therefore, it serves as a suitable
cultural context to investigate whether the claim made by Ji et al.
(2004) and previous studies -- that Eastern and Western cultures have
different notions of similarity -- extends beyond mainland China. In
addition to these replication and extension questions, we tested whether
fastText vectors from corpora corresponding to each language context
(English, Mandarin, and Vietnamese) are good predictors for similarity
judgments in each population.

This study is correlational and cannot evaluate causal relationships
between environmental statistics and similarity judgments. However, this
work can inform potential mechanisms by examining whether similarity
judgments co-vary with environmental statistics that differ across these
contexts.

Our specific research questions are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Do we replicate cross-cultural differences in similarity judgments
  between East Asian and Western cultures?
\item
  Are these cross-cultural differences related to differences in
  language?
\item
  Is our language model specific to items with a taxonomic-thematic
  contrast, or can it predict similarity more generally?
\end{enumerate}

To preview our results, we replicate US-China differences and find that
Vietnamese judgments are intermediate between these two. We find that
language-specific statistics provide good predictions for cross-cultural
differences in similarity judgments, and additionally, for more general
similarity judgments that do not contrast taxonomic and thematic
matches. Critically, the stimuli used to assess this more general case
of similarity reasoning were constructed to be outside the scope of
explanation for taxonomic-thematic accounts (for use as filler items).
While the language model may succeed in taxonomic-thematic similarity
predictions by picking up on differences in the \emph{conceptualization}
of similarity that are reflected in language data, this
conceptualization account provides no prediction for these filler
stimuli.

Taken together, our findings provide support for an alternative to
previous accounts on which differing cultures induce differing
conceptions of similarity. On this alternative, language explains for
culture-specific variation in similarity reasoning, both for
taxonomic-thematic judgments and for judgments not attributable to these
senses of similarity. While these findings do not rule out other
cross-cultural factors, they show it is possible to explain some
cross-cultural variation in similarity judgments purely from language
statistics.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

Data collection and analyses for this study were pre-registered, and the
pre-registration is available at {[}blinded for peer review{]}. We
recruited 200 participants from the US, 199 from Vietnam, and 200 from
mainland China. All participants were recruited through snowball
sampling, which, in the US was seeded with student email lists at a
large university, in Vietnam (VN) with Vietnam-based student groups on
Facebook, and in China (CN) with student groups on WeChat. US
participants were compensated with \$5 gift certificates (USD), VN
participants received 50,000₫ (VND) in phone credit, and CN participants
received 25CNY through WeChat credit transfer.

We excluded 8 US participants, 62 VN participants, and 16 CN
participants who missed two or more of the 10 attention check questions.

In doing so, we deviated from our preregistered exclusion criterion,
which used a more stringent policy, excluding participants who missed
any of the 10 attention checks. This approach would have substantially
reduced our sample size, disproportionately affecting the VN sample
(resulting in a final sample of 109 US, 132 CN, and 57 VN participants).
We report results with the less stringent exclusion criterion, but we
performed analyses with both samples and found broadly similar results.
Any differences are noted in the Analysis.

We also applied four demographic exclusion criteria designed to reduce
cross-cultural influences across our samples: we excluded participants
who (1) were not native speakers of the test language (English,
Vietnamese, or Mandarin Chinese), (2) were fluent in another one of the
study languages, (3) had lived outside of the test country for more than
two years, or (4) had significant international experience (more than 6
international experiences of 2 days or longer). Following our
pre-registration, we dropped any exclusion criterion that would exclude
25\% or more of any one population, but only for that population.
Specifically, we did not exclude VN and CN participants who spoke
English, or US participants with significant international experience.
As a result of these demographic exclusions, we excluded 73 US, 27 VN,
and 35 CN participants. The final US sample included 119 participants
(30M, 84F, 3 non-binary, 2 other), with mean age = 22.2 (SD = 8.15). Our
VN sample included 110 participants (34M, 71F, 5 other), with mean age =
22.21 (SD = 5.81). And the CN sample included 149 participants (61M,
87F, 1 other), with mean age = 23.1 (SD = 3.65).{[}\^{}4{]}

\hypertarget{materials}{%
\subsection{Materials}\label{materials}}

All stimuli and scripts for the experiment and analysis are available at
{[}BLINDED FOR REVIEW{]}. We collected and adapted stimuli from previous
studies to create test triads consisting of a cue with one thematic and
one taxonomic match option. For example, ``cow,'' ``grass,'' and
``chicken,'' where ``cow'' is the cue, ``grass'' is the thematic match,
and ``chicken'' the taxonomic match. We included 105 such triads, a
superset including triads pulled from supplemental information and
in-text examples across the literature, and others that we adapted or
created . We selected triads on the basis of cultural familiarity within
the US, Vietnam, and China. The triads were translated from English to
Vietnamese and Mandarin by fluent bilingual speakers of each language,
and back-translated to English by another fluent bilingual who was naive
to the original English version. The back-translations were checked
against the original English to identify ambiguities in the translation,
which were resolved through discussion and selection of alternative
terms when relevant.

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

Each participant completed all 105 triads in blocks of 21 trials at a
time (10 test triads, 10 filler triads, and 1 attention check per page),
by selecting the match most related to the cue (``Which thing is most
closely related to the bolded item?'' and translated equivalents). The
test triads were presented with 105 filler triads mixed in, to obscure
the taxonomic-thematic two-answer forced choice structure of the test
stimuli and reduce the likelihood that participants would become aware
of the design. The filler triads were groups of three semantically
related words, but where the match options were not distinguished by
thematic vs.~taxonomic similarity, for example, the cue ``bird'' with
match options ``lizard'' and ``toad.'' Additionally, we included 10
attention check trials, which were formatted similarly to the test and
filler triads but included an instruction instead of a cue item, e.g.,
``Choose wife'' with match options ``wife'' and ``husband.'' In total,
each participant completed 210 similarity judgments (105 test triads and
105 fillers) and 10 attention check questions, with all items presented
in randomized orders that varied between subjects.

\hypertarget{corpus-model}{%
\subsection{Corpus model}\label{corpus-model}}

Our general approach is to predict behavioral preferences in similarity
judgment using relative similarity between fastText word embeddings.

\hypertarget{word-vector-retrieval}{%
\paragraph{Word vector retrieval}\label{word-vector-retrieval}}

We use the fastText pre-trained models of English, Mandarin, and
Vietnamese in Grave, Bojanowski, Gupta, Joulin, \& Mikolov (2018). These
models were trained on Common Crawl and Wikipedia using fastText and use
character n-grams of length 5, and 10 negative examples. The training
used a Continuous Bag of Words with position-weights and a window of
size 5. From these models, we retrieve the word vectors (dimension 300)
for each word in our triad and filler stimuli.

\hypertarget{similarity-model}{%
\paragraph{Similarity model}\label{similarity-model}}

To give an intuition for our model, consider again the cow:
grass/chicken triad. We retrieved word vectors for ``cow'' and
``grass'', and calculated the cosine distance between these vectors. We
did the same for ``cow'' and ``chicken.'' Our similarity prediction is
inversely proportional to the ratio of cosine distance of these pairs.
This is because a larger cosine distance means the word vectors are
further apart, and thus the words are less similar. For example, if the
cosine distance of thematic cow-grass is 0.7 and the cosine distance of
taxonomic cow-chicken is 0.3 (the more similar of the two), then our
model predicts, correspondingly, that 30\% of responses to the triad
will be grass, and the other 70\% chicken.

We calculated the cosine distance between each cue-thematic match
(thematic cosine distance) and cue-taxonomic match (taxonomic cosine
distance). We then calculated the thematic cosine distance proportion as
thematic cosine distance over the sum of taxonomic cosine distance and
thematic cosine distance. We followed this process for all three
corpora, and were able to obtain predictions for all triads in all
languages. Then, we used a mixed-effects regression to evaluate how well
each corpus model predicts participants' similarity judgments, across
triads and cultural contexts.

\hypertarget{results-discussion}{%
\section{Results \& Discussion}\label{results-discussion}}

\hypertarget{replication-of-previous-work-and-extension-to-a-vietnamese-sample}{%
\subsection{1. Replication of previous work and extension to a
Vietnamese
sample}\label{replication-of-previous-work-and-extension-to-a-vietnamese-sample}}

In keeping with Ji et al. (2004), we would expect participants from
mainland China to prefer thematic matches more than US participants
(e.g., to prefer the cow-grass match over cow-chicken to a greater
extent than US participants). On the basis of the previous literate more
broadly, we would also expect participants from Vietnam to pattern with
China, showing a stronger preference for thematic matches than US
participants.

We observed the strongest preference for thematic matching in China (M =
0.65, SD = 0.11), followed by Vietnam (M = 0.6, SD = 0.11), and then the
US (M = 0.56, SD = 0.17; see Figure 1).

\begin{CodeChunk}
\begin{figure}[tb]

{\centering \includegraphics{figs/unnamed-chunk-1-1}

}

\caption[Proportion of thematic responses by country]{Proportion of thematic responses by country.}\label{fig:unnamed-chunk-1}
\end{figure}
\end{CodeChunk}

To test for differences in similarity judgment between the countries, we
ran a mixed-effects logistic regression predicting triad match
(taxonomic or thematic) with country (US, China, or Vietnam) as a fixed
effect. As random effects, we included an intercept per subject and one
per triad, as well as by-triad random slopes for country to account for
variation in the country effect across triads.

We found a significant effect of country (\(\chi^2\)(2) = 15.37, p
\textless{} .001), but this effect is driven by the difference between
US and CN responding (\(\beta\) = -0.48, p \textless{} .001). There was
no statistical difference between matching in Vietnam and China
(\(\beta\) = -0.22, p = 0.09), or the US and Vietnam (\(\beta\) = 0.26,
p = 0.086).

In sum, we replicated the differences documented by Ji et al. (2004)
between the US and China. We did not find that Vietnamese participants
grouped particularly with Chinese participants, however. Instead they
were intermediate between the two. Accordingly, we did not find strong
support for the hypothesis that there are overall biases toward thematic
responding across Asian cultural contexts broadly.

\hypertarget{language-statistics-as-a-predictor-for-cross-cultural-variation-in-similarity-judgments}{%
\subsection{2. Language statistics as a predictor for cross-cultural
variation in similarity
judgments}\label{language-statistics-as-a-predictor-for-cross-cultural-variation-in-similarity-judgments}}

In our next analyses, we tested whether triad-level differences in
judgments across countries could be predicted on the basis of language
statistics.

\hypertarget{single-corpus-model}{%
\subsubsection{Single corpus model}\label{single-corpus-model}}

First, to test whether variation in language statistics can explain
differences in similarity judgments, we use a mixed-effects logistic
regressions fit individually for each country. Our models predicted
responses by individual participants to particular triads (0=taxonomic
or 1=thematic) with fastText predictions (proportion of cosine distance)
as a fixed effect and participant and triad as random effects. If
language statistics, contribute to the differences in similarity
judgments, we would expect each language corpus to be a good predictor
for similarity judgment responding in its corresponding context.

All corpora were significant predictors of all cultural context
responding, with p \textless{} 0.05 and \(\beta\) from -8.69 to -2.4.

\hypertarget{multi-corpus-model}{%
\subsubsection{Multi-corpus model}\label{multi-corpus-model}}

If language statistics are able to predict meaningful culture-specific
variation in similarity judgments (rather than just consistency across
cultures), we would expect each corpus to be the best predictor of its
corresponding culture compared to the other two corpora. We directly
compared the corpus models by including all three corpus predictors as
fixed effects in three mixed-effect regressions (predicting US, Vietnam
and China responding) with the same random effects as above.

For US responding: only the English (EN) corpus was a significant
predictor\footnote{With the preregistered exclusion criterion, only the
  English (EN) and Mandarin (ZH) corpus were significant predictors.}.
EN corpus: \(\beta\) = -6.96, \(\chi^2\)(1) = 16.57, p \textless{} .001.
VI corpus: \(\beta\) = -2.22, \(\chi^2\)(1) = 3.73, p = 0.054. ZH
corpus: \(\beta\) = -3.18, \(\chi^2\)(1) = 3.37, p = 0.066.

For Vietnam responding, only the Vietnamese (VI) and Mandarin (ZH)
corpus were significant predictors\footnote{With the preregistered
  exclusion criterion, all three corpora were significant predictors.}.
EN corpus: \(\beta\) = -2.98, \(\chi^2\)(1) = 2.58, p = 0.108. VI
corpus: \(\beta\) = -2.75, \(\chi^2\)(1) = 4.84, p = 0.028. ZH corpus:
\(\beta\) = -4.39, \(\chi^2\)(1) = 5.49, p = 0.019.

For China responding, only the Mandarin (ZH) and English (EN) corpus
were significant predictors. EN corpus: \(\beta\) = -3.32, \(\chi^2\)(1)
= 7.18, p = 0.007. VI corpus: \(\beta\) = -1.59, \(\chi^2\)(1) = 3.63, p
= 0.057. ZH corpus: \(\beta\) = -5.31, \(\chi^2\)(1) = 17.69, p
\textless{} .001.

We observed some language specificity in this analysis (Figure 2). The
English corpus was the best predictor for US responding, and the
Mandarin corpus was the best predictor for Chinese responding. While
this is not the case with the Vietnamese corpus and the Vietnam
responding, the Vietnamese corpus was still a significant predictor for
the Vietnam responding. These results support our hypothesis that
specific language statistics can predict cross-cultural variations of
similarity judgment.

\begin{CodeChunk}
\begin{figure}[tb]

{\centering \includegraphics{figs/coeffs_cos-1}

}

\caption[Fixed effect sizes of each corpus lexical statistics (cosine distance proportion) when included as a predictor for China, Vietnam, and US responding, respectively]{Fixed effect sizes of each corpus lexical statistics (cosine distance proportion) when included as a predictor for China, Vietnam, and US responding, respectively. The English corpus is the best predictor for US response, and the Mandarin corpus is the best predictor for China response.}\label{fig:coeffs_cos}
\end{figure}
\end{CodeChunk}

However, in all cultural contexts, adding the other two corpora produced
a significantly better fit than the identical model with only the
corresponding corpus included as a predictor (US response: \(\chi^2\)(2)
= 7.93, p = 0.019; Vietnam response: \(\chi^2\)(2) = 13.72, p = 0.001;
China response: \(\chi^2\)(2) = 10.56, p = 0.005). This analysis
suggests that culture-specific inputs to similarity judgment (as proxied
by language statistics) do not fully explain cross-cultural differences
in similarity judgment, whether due to noise in the estimates of
language statistics or true additional variance picked up by embeddings
from other corpora.

\hypertarget{language-statistics-as-a-predictor-for-non-structured-filler-items}{%
\subsection{3. Language statistics as a predictor for non-structured
filler
items}\label{language-statistics-as-a-predictor-for-non-structured-filler-items}}

One possible concern with our approach is that our models are only able
to account for variation in similarity judgment because the word
embedding models pick up on cross-cultural differences in the
conceptualization of similarity. Perhaps some aspect of the geometry of
the embedding space specifically reflects these differences rather than
capturing more item-specific patterns of co-occurrence.

If this is the case, the embedding models should be less accurate at
predicting responding for filler items (such as whether ``tomorrow'' or
``yesterday'' is more related to ``today'') because these items do not
have a thematic/taxonomic structure. Additionally, even when one option
might be more related to the cue, that relationship is not systematic
throughout the set of fillers. On the other hand, if the embedding
models capture similarity judgments more generally (beyond
thematic/taxonomic preferences), they should predict judgments for the
unstructured filler items as well as they do for our experimental items.

For each filler triad, we randomly assigned one of the responding
options as `Word1'. Running a mixed-effects logistic regression to
predict responding (Word1 or Word2) with country as a fixed effect and a
random effect structure as above Q1, we found no effect of country on
filler responding (\(\chi^2\)(2) = 0.74, p = 0.692).

Using the same mixed-effects logistic regression structure as the single
corpus models in Q2, we predicted responses (1=word1 or 0=word2) in each
cultural context with the corresponding corpus predictor as a fixed
effect, and participant and triad as random effects. In all cultural
contexts, the corresponding corpus was a significant predictor of
responding, with p \textless{} 0.05 and \(\beta\) from -9.59 to -3.14.
These results show that the word embedding models predict general
similarity in addition to structured thematic/taxonomic similarity.

To investigate whether language statistics predict taxonomic/thematic
triads and filler items differently, we compared the variance accounted
for by the models of the two different types of items. We ran two
mixed-effects logistic regression models that predict responding across
cultural contexts (1=thematic or word1, 0=taxonomic or word2), with the
corresponding corpus predictor as a fixed effect and participant and
item as random effects. Consistent with our results above, corresponding
corpus was a significant predictor for responding to both triads and
filler items (triads: \(\beta\) = -1.79, \(\chi^2\)(1) = 75.58, p
\textless{} .001, fillers: \(\beta\) = -2, \(\chi^2\)(1) = 70.89, p
\textless{} .001). Importantly, we found comparable conditional \(R^2\)
values when the corresponding corpus was used to predict only triad
items (\(R^2\) = 0.3) and only filler items (\(R^2\) = 0.26). This
result provides evidence for our view that word embeddings index general
tendencies in similarity judgment, beyond a specific dimension of
thematic/taxonomic relations.

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

Do members of different cultures vary in their conceptualization of
similarity? Cross-cultural differences in similarity judgment would seem
to suggest this conclusion. In this paper, we consider whether
statistics of the environment (as indexed by language statistics) can
account for cross-cultural differences in a classic similarity judgment
paradigm. We replicated the previously documented contrast between
English speakers in the US and Mandarin Chinese speakers from East Asia
(mainland China, Taiwan, Hong Kong, and Singapore): Chinese participants
preferred thematic relations to a greater extent compared to US
participants. Our sample of Vietnamese participants showed an
intermediate preference but were not significantly different from either
Chinese or US participants. This finding suggests some limitations on
the generality of the cultural account, which proposes that
thematic/taxonomic similarity preference aligns with East Asian/Western
tendency for holistic vs analytic processing.

In contrast, we found some support for the environmental statistics
account: word embeddings based on language specific corpora were a good
predictor for the corresponding country's similarity judgments, even
when other corpus statistics were included, and even with filler triads
without a thematic/taxonomic structure. Overall, our results provide
evidence that cross-cultural differences in similarity judgment are
related to linguistic statistics patterns (which also vary across
cultures).

There are some important limitations of our approach. While we discuss
cross-cultural variability at the level of countries or larger world
areas, these are not cultural monoliths. For convenience, we
operationalize culture at the level of country, based on where
participants were raised. It is an open question whether performance in
our participant populations (of relatively young and well-educated
adults) is representative of the broader country. This is especially
true for societies with substantial ethnic and cultural variation such
as the US. We expect that our data is likely to underestimate variation
both within and between the countries we sample from.

Additionally, language, culture, cognition, and individual experiences
are intertwined in complex causal relationships. In this study, we
measure language and its relation to cross-cultural differences in
categorization, but these relations test only the plausibility of a
language-based account; they cannot establish the direction of
causality.

Ji et al. (2004) established that culture-aligned differences in this
paradigm exist, even when the test language is held constant, concluding
that ``it is culture (independent of the testing language) that led to
different grouping styles'' in their study. Our data provide a
cautionary note to this conclusion, suggesting that while cultural
differences in similarity judgment exists, we might get better traction
on modelling cross-cultural variation through environmental statistics
such as lexical co-occurrences.

There are still many open questions for this account, however. Our
operationalization of environmental statistics with word embeddings
likely captures both language inputs and culture-specific
conceptualization of similarity. Further investigation of the semantic
spaces captured by cross-cultural word embedding models is required.
Future work should also aim to provide a more specific computational
account of how lexical co-occurrence might guide categorization
preference beyond the simple proportion-of-similarity model tested here,
and investigate other individual- or triad-specific factors that might
influence how environmental statistics affect similarity judgment.

Despite these caveats, our findings here demonstrate the plausibility of
an alternative perspective on cross-cultural accounts of similarity in
the case of taxonomic and thematic reasoning. It may be the input to
similarity judgments, rather than the evaluative process or the
conceptualization of similarity that produces variation in similarity
reasoning across cultural and linguistic contexts. We hope this work
provides a foundation for further research probing this question.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

BLINDED FOR PEER REVIEW.

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}

\noindent

\newpage

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-asr2018}{}}%
Asr, F. T., Zinkov, R., \& Jones, M. (2018). Querying word embeddings
for similarity and relatedness. In \emph{Proceedings of the 2018
conference of the north american chapter of the association for
computational linguistics: Human language technologies, volume 1 (long
papers)} (pp. 675--684).

\leavevmode\vadjust pre{\hypertarget{ref-Chiu1972}{}}%
Chiu, L. (1972). A cross-cultural comparison of cognitive styles in
chinese and american children. \emph{International Journal of
Psychology}, \emph{7}(4), 235--242.
http://doi.org/\href{https://doi.org/doi:10.1080/00207597208246604}{doi:10.1080/00207597208246604}

\leavevmode\vadjust pre{\hypertarget{ref-Choi1999}{}}%
Choi, I., Nisbett, R. E., \& Norenzayan, A. (1999). Causal attribution
across cultures: Variation and universality. \emph{International Journal
of Psychology}, \emph{125}(1), 47--63.
http://doi.org/\href{https://doi.org/doi:10.1037/0033-2909.125.1.47}{doi:10.1037/0033-2909.125.1.47}

\leavevmode\vadjust pre{\hypertarget{ref-Chua2005}{}}%
Chua, H. F., Boland, J. E., \& Nisbett, R. E. (2005). Cultural variation
in eye movements during scene perception. \emph{Proceedings of the
National Academy of Sciences}, \emph{102}(35), 12629--12633.
http://doi.org/\href{https://doi.org/10.1073/pnas.0506162102}{10.1073/pnas.0506162102}

\leavevmode\vadjust pre{\hypertarget{ref-Grave2018}{}}%
Grave, E., Bojanowski, P., Gupta, P., Joulin, A., \& Mikolov, T. (2018).
Learning word vectors for 157 languages. In \emph{Proceedings of the
international conference on language resources and evaluation (LREC
2018)}.

\leavevmode\vadjust pre{\hypertarget{ref-Griffiths2007}{}}%
Griffiths, T. L., Steyvers, M., \& Tenenbaum, J. B. (2007). Topics in
semantic representation. \emph{Psychological Review}, \emph{114}(2),
211--244.
http://doi.org/\href{https://doi.org/doi:10.1037/0033-295x.114.2.211}{doi:10.1037/0033-295x.114.2.211}

\leavevmode\vadjust pre{\hypertarget{ref-Hui2002}{}}%
Hui, W. (2002). Modernity and 'asia' in the study of chinese history. In
E. Fuchs \& B. Stuchteyi (Eds.), \emph{Across cultural borders:
Historiography in global perspective}. Rowman \& Littlefield.

\leavevmode\vadjust pre{\hypertarget{ref-Jatnika2019}{}}%
Jatnika, D., Bijaksana, M. A., \& Suryani, A. A. (2019). Word2Vec model
analysis for semantic similarities in english words. \emph{Procedia
Computer Science}, \emph{157}, 160--167.
http://doi.org/\url{https://doi.org/10.1016/j.procs.2019.08.153}

\leavevmode\vadjust pre{\hypertarget{ref-Ji2000}{}}%
Ji, L., Peng, K., \& Nisbett, R. E. (2000). Culture, control, and
perception of relationships in the environment. \emph{Journal of
Personality and Social Psychology}, \emph{78}(5), 943--955.
http://doi.org/\href{https://doi.org/doi:10.1037/0022-3514.78.5.943}{doi:10.1037/0022-3514.78.5.943}

\leavevmode\vadjust pre{\hypertarget{ref-Ji2004}{}}%
Ji, L., Zhang, Z., \& Nisbett, R. E. (2004). Is it culture or is it
language? Examination of language effects in cross-cultural research on
categorization. \emph{Journal of Personality and Social Psychology},
\emph{87}(1), 57--65.
http://doi.org/\href{https://doi.org/doi:10.1037/0022-3514.87.1.57}{doi:10.1037/0022-3514.87.1.57}

\leavevmode\vadjust pre{\hypertarget{ref-Liu2019}{}}%
Liu, N., Feng, C., Wu, S., Chan, A., \& Fulton, J. (2019). Automate RFP
response generation process using FastText word embeddings and soft
cosine measure. In \emph{Proceedings of the 2019 international
conference on artificial intelligence and computer science} (pp.
12--17).
http://doi.org/\href{https://doi.org/10.1145/3349341.3349362}{10.1145/3349341.3349362}

\leavevmode\vadjust pre{\hypertarget{ref-Markman1984}{}}%
Markman, E. M., \& Hutchinson, J. E. (1984). Children's sensitivity to
constraints on word meaning: Taxonomic versus thematic relations.
\emph{Cognitive Psychology}, \emph{16}(1), 1--27.
http://doi.org/\href{https://doi.org/doi:10.1016/0010-0285(84)90002-1}{doi:10.1016/0010-0285(84)90002-1}

\leavevmode\vadjust pre{\hypertarget{ref-Masuda2001}{}}%
Masuda, T., \& Nisbett, R. E. (2001). Attending holistically versus
analytically: Comparing the context sensitivity of japanese and
americans. \emph{Journal of Personality and Social Psychology},
\emph{81}(5), 922--934.
http://doi.org/\href{https://doi.org/doi:10.1037/0022-3514.81.5.922}{doi:10.1037/0022-3514.81.5.922}

\leavevmode\vadjust pre{\hypertarget{ref-Mikolov2018}{}}%
Mikolov, T., Grave, E., Bojanowski, P., Puhrsch, C., \& Joulin, A.
(2018). Advances in pre-training distributed word representations. In
\emph{Proceedings of the international conference on language resources
and evaluation (LREC 2018)}.

\leavevmode\vadjust pre{\hypertarget{ref-Nisbett2003}{}}%
Nisbett, R. E. (2003). \emph{The geography of thought: How asians and
westerners think differently {\ldots{}} and why}. New York: Free Press.

\leavevmode\vadjust pre{\hypertarget{ref-Nisbett2003b}{}}%
Nisbett, R. E., \& Masada, T. (2003). Culture and point of view.
\emph{Proceedings of the National Academy of Sciences}, \emph{100}(19),
11163--11170.
http://doi.org/\href{https://doi.org/10.1073/pnas.1934527100}{10.1073/pnas.1934527100}

\leavevmode\vadjust pre{\hypertarget{ref-Norenzayan2002}{}}%
Norenzayan, A., Smith, E. E., Kim, B. J., \& Nisbett, R. E. (2002).
Cultural preferences for formal versus intuitive reasoning.
\emph{Cognitive Science}, \emph{26}(5), 653--684.
http://doi.org/\href{https://doi.org/doi:10.1207/s15516709cog2605_4}{doi:10.1207/s15516709cog2605\_4}

\leavevmode\vadjust pre{\hypertarget{ref-Rohde2006}{}}%
Rohde, D. L., Gonnerman, L. M., \& Plaut, D. C. (2006). An improved
model of semantic similarity based on lexical co‐occurrence.
\emph{Communications of the ACM}, \emph{8}, 627--633.

\leavevmode\vadjust pre{\hypertarget{ref-Thompson2020}{}}%
Thompson, R., B., \& Lupyan, G. (2020). Cultural influences on word
meanings revealed through large-scale semantic alignment. \emph{Nature
Human Behaviour Volume}, \emph{4}, 1029--1038.
http://doi.org/\url{https://doi.org/10.1038/s41562-020-0924-8}

\end{CSLReferences}

\bibliographystyle{apacite}

\end{document}